---
title: "(Gentle) Cluster Analysis of the World by Happiness"
output: html_notebook
---

```{r echo = F, message = F}
library(tidyverse)
library(data.table)
library(DT)
```

```{r}
wHappiness_Df <- read.csv("C:/Users/PC/Documents/R_4DS/WorldHappinessReport/world-happiness-report-2021.csv")
glimpse(wHappiness_Df)
```

## Data Inspection and Cleaning

```{r}
## Check for missing values
null_vars <- (sapply(wHappiness_Df, function(x) sum(is.na(x))))
t(data.frame(null_vars))

## Checking for Blank Values
wHappiness_Df[wHappiness_Df == "",]
```

No missing or empty values, go-to-go.

```{r}
names(wHappiness_Df)
```


```{r}
## Tidying the Column Name
colClean <- function(x){
  colnames(x) <- gsub("\\.\\.+",".", colnames(x));
  x
}

wHappiness_Df <- colClean(wHappiness_Df) %>% 
  rename(Country = ï.Country.name)
names(wHappiness_Df)
```

```{r}
# colnames(wHappiness_Df) <- gsub("Explained.by.", "", colnames(wHappiness_Df))
# names(wHappiness_Df)
```


## Exploratory Data Analysis

```{r}
wHappiness_Df %>% 
  count(Regional.indicator) %>% 
  ggplot(aes(x = reorder(Regional.indicator, n), y = n)) +
  geom_bar(stat = "identity") +
  coord_flip()
```

#### Regions by LifeExpectancy
```{r}
# plotdata <- function(xx, y){
#   if(is.integer(y) == TRUE){
#     wHappiness_Df %>% 
#     group_by(xx) %>% 
#     summarise(mean_ = mean(y))
#   } else {
#     print("Enter Integer")
#   }
# }
# 
# plotdata(Regional.indicator, Healthy.life.expectancy)

plotdata <- wHappiness_Df %>%
  group_by(Regional.indicator) %>%
  summarise(n = n(),
            mean_ = mean(Healthy.life.expectancy),
            sd = sd(Healthy.life.expectancy),
            se = sd/sqrt(n),
            ci = qt(0.0975, df = n - 1) * sd/sqrt(n))

plotdata %>% 
  ggplot(aes(x = reorder(Regional.indicator, mean_), y = mean_, group = 1)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = mean_ - se,
                    ymax = mean_ + se),
                width = .1) +
  coord_flip()
```


## Cluster Analysis

```{r}
cluster_x <- wHappiness_Df %>% 
  select(-c("Regional.indicator", "Country"))

##
cluster_x <- cluster_x[ , which(apply(cluster_x, 2, var) != 0)]
which(apply(cluster_x, 2, var)==0)

## Checking for clusters
library(factoextra)
library(gridExtra)

set.seed(234)

## Initiate clusters with k = (3,4,5,6)
k0 <- kmeans(cluster_x, centers = 2, nstart = 25)
k1 <- kmeans(cluster_x, centers = 3, nstart = 25)
k2 <- kmeans(cluster_x, centers = 4, nstart = 25)
k3 <- kmeans(cluster_x, centers = 5, nstart = 25)
k4 <- kmeans(cluster_x, centers = 6, nstart = 25)

##Visualize cluster results
p0 <- fviz_cluster(k0, geom = "point", data = cluster_x) + ggtitle("K means k=3")
p1 <- fviz_cluster(k1, geom = "point", data = cluster_x) + ggtitle("K means k=3")
p2 <- fviz_cluster(k2, geom = "point", data = cluster_x) + ggtitle("K means k=4")
p3 <- fviz_cluster(k3, geom = "point", data = cluster_x) + ggtitle("K means k=5")
p4 <- fviz_cluster(k4, geom = "point", data = cluster_x) + ggtitle("K means k=6")


## Concanate grids
grid.arrange(p0, p1, p2, p3, p4)
```

```{r}
## Determine n of cluster, plot Elbow method
fviz_nbclust(cluster_x, kmeans, method = "wss") +
  geom_vline(xintercept = 2, linetype = 2) 
```

k = 2 forms the perfect elbow


## Evaluation

```{r}
## Applying n = 4 clusters on dataset.
cluster_Df <- data.frame(wHappiness_Df, Cluster = k0$cluster) %>% 
  mutate(Cluster = as.factor(Cluster))

cluster_Df %>% 
  group_by(Cluster) %>% 
  ggplot(aes(x = Regional.indicator, y = Healthy.life.expectancy, colour = Cluster)) +
  geom_jitter()  +
  facet_wrap(~Cluster) +
  coord_flip()
  
```

#### Ladder score of Happiness

```{r}
cluster_Df %>% 
  group_by(Cluster) %>% 
  ggplot(aes(x = Regional.indicator, y = Ladder.score, colour = Cluster)) +
  geom_jitter()  +
  facet_wrap(~Cluster) +
  coord_flip()
```



#### Logged GDP/Capita

```{r}
cluster_Df %>% 
  group_by(Cluster) %>% 
  ggplot(aes(x = Regional.indicator, y = Logged.GDP.per.capita, colour = Cluster)) +
  geom_jitter()  +
  facet_wrap(~Cluster) +
  coord_flip()
```

#### Perception of Corruption
```{r}
cluster_Df %>% 
  group_by(Cluster) %>% 
  ggplot(aes(x = Regional.indicator, y = Perceptions.of.corruption, colour = Cluster)) +
  geom_jitter()  +
  facet_wrap(~Cluster) +
  coord_flip()
```


Perception of corruption in the Second Cluster is more generalized, unlike the first cluster who have a more defined and uniform perception of corruption, however provided by the survey's scale.

At this junction our evaluation of our cluster analysis gives to define the clusters in pre-concieved real world ideas of (k = 1)Developed and (k =2) Developing Clusters/Countries, which are found, ironically, to rate higher on the scale of happiness. Indicatively more has to be done to attenuate the real reasons behind these metrics.

```{r}
cluster_Df <- cluster_Df %>% 
  mutate(Cluster = if_else(Cluster == 1, "Developed_Region", "Developing_Region"))
```

